---
title: "Project 9"
author: "CodesRUS"
date: "12/13/2021"
output: html_document
bibliography: BIOL3140.bib
---


### Introduction
	Trans-gulf migrants, known as TGMs, are species that endure the long migration over the Gulf of Mexico from Central and South America before arriving on land between Texas and Florida. The birds usually migrate during the springtime in a nonstop flight in hopes to find newer land in which they can forage and breed on. The specific time that the birds arrive on the North American land can have major impacts on the fitness of the bird overall. Birds who arrive sooner tend to have better opportunities to find a mate, whereas birds that arrive later have a more ensured food stability. In response to climate change and warming temperatures, TGMs need to shift the date in which they arrive at their breeding grounds or we could see mass population declines. Scientists are currently debating the extent to which TGMs and other long-distance migrants are shifting their arrival patterns, if they are at all. In order to determine the magnitude of these shifts, scientists have been using field, tracking, and banding methods but data has been limited as there are not many scientists who do this type of work.
	In more recent years, trends over species conservation and climate change awareness have inspired further research and developments in similar issues. One such study conducted by (@charmantier2008adaptive) highlighted the importance of understanding the mechanisms behind population responses to climate change. Phenotypic plasticity is one such mechanism they studied and found that individual responses to these environmental changes may be fixed in the population but evolve over time. Specifically concerning TGMs, many online databases have arisen that allow you to log observations about birds that can be seen by the public. One in particular, fittingly titled eBird, is one of the most complex and has proven to be especially helpful to scientists. Created by the Cornell Laboratory of Ornithology and the National Audubon Society, eBird launched in 2002 and has over 100 million bird observations made by thousands of different people. The database even allows observers to log historical data so observations from pre-2002 can be read as well. In this project we want to use eBird and meteorological data to study what affects the local weather conditions can have on the arrival times of TGMs in the state of Massachusetts.

### Methods
In order to find the effect of local weather conditions on the arrival time of TGMs in Massachusetts, we first use f loop to load all the data of human observation in Massachusetts from eBird database. Then, we use a token obtained from NOAA website to download weather data of three sample locations. After that, we move to the data analysis section. To prepare the eBird data, we plot the proportion of the population that has arrived and predict their arrival time. At the end of this part, we plot how this arrival day varies with year to see the influence of climate on arrival date. To prepare for weather data, we calculate wind direction and join the weather data with species. Next, we calculate the mean of our weather variables 5 and 10 days before by using frollmean() function to get more accurate information. Finally, we use lme from lme4 package to produce Linear Mixed-effect Model. Anova test is performed to find the best fit model and dredg() function is used for model testing.

### Results

```{r, include=FALSE}
library(tidyverse)
library(MuMIn)
library(rnoaa)
library(data.table)
library(ggmap)
library(usmap)
library(magick)
library(cowplot)
library(lme4) 
library(car) 
library(data.table) 
library(rgbif)
library(rgdal)
library(knitr)
library(kableExtra)
library(systemfonts)
```

```{r, "species images & data", warning=FALSE, message=FALSE}
### choice of species 
#Chimney Swift, Chaetura pelagica
#Ruby-throated Hummingbird, Archilochus colubris
#Belted Kingfisher, Megaceryle alcyon
#Yellow-bellied Sapsucker, Sphyrapicus varius
#Common Yellowthroat,Geothlypis trichas


swift <- occ_data(scientificName = "Chaetura pelagica", stateProvince = "Massachusetts", limit = 200, year = 2018)
humming<- occ_data(scientificName = "Archilochus colubris", stateProvince = "Massachusetts", limit = 200, year = 2018)
kingfisher<- occ_data(scientificName = "Megaceryle alcyon", stateProvince = "Massachusetts", limit = 200, year = 2018)
sapsucker<- occ_data(scientificName = "Sphyrapicus varius", stateProvince = "Massachusetts", limit = 200, year = 2018)
yellowthroat<- occ_data(scientificName = "Geothlypis trichas", stateProvince = "Massachusetts", limit = 200, year = 2018)

MA <- map_data('state', 'massachusetts')

swift.p <- ggplot(MA, aes(long,lat,group=subregion) )+
  geom_polygon(colour = "gray",fill="gray90")+geom_point(data=swift[[2]],aes(x=decimalLongitude,y=decimalLatitude,size=individualCount),alpha=0.3,inherit.aes = F)+ coord_quickmap()+theme_void()
humming.p <- ggplot(MA, aes(long,lat,group=subregion) )+
  geom_polygon(colour = "gray",fill="gray90")+geom_point(data=humming[[2]],aes(x=decimalLongitude,y=decimalLatitude,size=individualCount),alpha=0.3,inherit.aes = F)+ coord_quickmap()+theme_void()
kingfisher.p <- ggplot(MA, aes(long,lat,group=subregion) )+
  geom_polygon(colour = "gray",fill="gray90")+geom_point(data=kingfisher[[2]],aes(x=decimalLongitude,y=decimalLatitude,size=individualCount),alpha=0.3,inherit.aes = F)+ coord_quickmap()+theme_void()
sapsucker.p <- ggplot(MA, aes(long,lat,group=subregion) )+
  geom_polygon(colour = "gray",fill="gray90")+geom_point(data=sapsucker[[2]],aes(x=decimalLongitude,y=decimalLatitude,size=individualCount),alpha=0.3,inherit.aes = F)+ coord_quickmap()+theme_void()
yellowthroat.p <- ggplot(MA, aes(long,lat,group=subregion) )+
  geom_polygon(colour = "gray",fill="gray90")+geom_point(data=yellowthroat[[2]],aes(x=decimalLongitude,y=decimalLatitude,size=individualCount),alpha=0.3,inherit.aes = F)+ coord_quickmap()+theme_void()

swift.p2 <- ggdraw() + draw_image("swift.png", scale =0.3, halign=0, valign=1) + draw_plot(swift.p)
print(swift.p2)
humming.p2 <- ggdraw() + draw_image("humming.png", scale =0.3, halign=0, valign=1) + draw_plot(humming.p)
print(humming.p2)
kingfisher.p2 <- ggdraw() + draw_image("kingfisher.png", scale =0.3, halign=0, valign=1) + draw_plot(kingfisher.p)
print(kingfisher.p2)
sapsucker.p2 <- ggdraw() + draw_image("sapsucker.png", scale =0.3, halign=0, valign=1) + draw_plot(sapsucker.p)
print(sapsucker.p2)
yellowthroat.p2 <- ggdraw() + draw_image("yellowthroat.png", scale =0.3, halign=0, valign=1) + draw_plot(yellowthroat.p)
print(yellowthroat.p2)
```
The above are the records for each of the 5 species selected in Massachusetts, 2018.

```{r, "loading data and creating proper data set", echo=TRUE, results=FALSE, message=FALSE}
species <- c("Chaetura pelagica","Archilochus colubris","Megaceryle alcyon","Sphyrapicus varius","Geothlypis trichas")

y <- paste0("2000",",","2019")
m <- paste0("4",",","5")

dat.l <-list()
for(s in species){
  n.obs <-  occ_data(scientificName = s,year=y,month=m,limit=0,country="US",basisOfRecord = "HUMAN_OBSERVATION",stateProvince="Massachusetts")$meta$count 
  
  print(n.obs)
  
  dat.l[[paste0(s)]] <- occ_data(scientificName = s,year=y,month=m,
                                 limit=n.obs,country="US",
                                 basisOfRecord = "HUMAN_OBSERVATION",
                                 stateProvince="Massachusetts")[[2]]
  
}

dat <- rbindlist(dat.l,fill=T)

head(dat)

saveRDS(dat,"massbird.data.RDS")

library(tidyverse)
dat <- readRDS("massbird.data.RDS")

dat.l%>%
  bind_rows(.id = "species") %>% 
  group_by(year,species)%>%
  summarise(count=sum(individualCount,na.rm = T))%>%
  ggplot(aes(x=year,y=count,col=species))+geom_point()
```
To begin our analysis, collection and sorting of data needed to occur for each species selected within a specific state during a range of years. A data set was created to do such.

```{r, "querying NOAA's NCDC API", warning=FALSE, results=FALSE}
options(noaakey = "VYpNYgvfxWWfUIWmcjVZxUUJwebzTmLB")

sts <- c(
  "GHCND:USW00013894", #Mobible, AL 2k away about 10 days away @200 km/day
  "GHCND:USW00013881", #Charlotte, NC 1000 km away about 6 days away @200 km/day
  "GHCND:USW00014739" #Boston
)

bos <- ncdc_stations(stationid = "GHCND:USW00014739")
print(bos)

cha <- ncdc_stations(stationid = "GHCND:USW00013881")
print(cha)

mob <- ncdc_stations(stationid = "GHCND:USW00013894")
print(mob)


sta.d <- bind_rows( #bind the rows
  lapply(sts,function(x) ncdc_stations(stationid = x)$data ) 
)%>%
  left_join(usmap_transform(.[,c("longitude","latitude")]))%>% 
  mutate(name=str_sub(name, -5,-4))%>%
  mutate(migr.day=c(10,5,0))%>% 
  separate(id,into = c("station.type","id"))%>%
  print()

plot_usmap(
  include = c(.northeast_region,.south_region,.east_north_central)
)+geom_point(data=sta.d,aes(x=longitude.1,y=latitude.1,col=name),size=5)+geom_label(data=sta.d,aes(x=longitude.1,y=latitude.1,col=name,label=name),size=5,nudge_x = 1e6*0.25)+theme(legend.position = "none")

weather.d <- meteo_pull_monitors(sta.d$id,date_min = "2005-01-01") #since 2005 we see an uptick in eBird data

head(weather.d)
```
Using NOAA's NCDC API, weather parameters during certain time frames were collected and placed within data sets. Weather sampling locations for this project are represented above. 

```{r, "eBird data for chimney swift (cs)", warning=FALSE, results=FALSE}
cs<- dat.l%>%
  bind_rows(.id = "species") %>% 
  filter(species=="Chaetura pelagica")%>%
  group_by(year)%>%
  mutate(date=as.Date(paste0(year,"-",month,"-",day)),
         j.day=julian(date,origin=as.Date(paste0(unique(year),"-01-01")))
  )%>%
  group_by(species,year,j.day,date)%>%
  summarise(day.tot=sum(individualCount,na.rm=T))%>%
  group_by(species,year)%>%
  mutate(prop=cumsum(day.tot/sum(day.tot,na.rm = T)))%>%
  filter(year>2004)

cs%>%
  ggplot(aes(j.day,prop))+geom_point()+facet_wrap(year~.)

cs.pred <- cs%>%
  group_by(year)%>%
  summarize(
    pred=predict(nls(prop~SSlogis(j.day,Asym, xmid, scal)),newdata=data.frame(j.day=min(j.day):max(j.day))), 
    j.day=min(j.day):max(j.day),
  )%>%
  left_join(cs%>%select(j.day,date)) 

cs%>%
  ggplot(aes(j.day,prop))+geom_point(aes=0.3)+geom_line(data=cs.pred,aes(x=j.day,y=pred),col="blue",size=2)+facet_wrap(year~.)

cs.arrive.date <-cs.pred%>%
  group_by(year)%>%
  filter(j.day==j.day[which.min(abs(pred-0.25))])

cs.arrive.date%>%
  ggplot(aes(year,j.day))+geom_point()
```
For Chimney Swift, the arrivals are depicted above with the estimation of arrivals. Then, using this data we plotted how the arrival day varied depending on the year. Based on the plots, the Julian day when 25% of the population arrived were 127-130.

```{r, "eBird data for ruby-throated hummingbird", warning=FALSE, results=FALSE}
rh<- dat.l%>%
  bind_rows(.id = "species") %>% 
  filter(species=="Archilochus colubris")%>%
  group_by(year)%>%
  mutate(date=as.Date(paste0(year,"-",month,"-",day)),
         j.day=julian(date,origin=as.Date(paste0(unique(year),"-01-01")))
  )%>%
  group_by(species,year,j.day,date)%>%
  summarise(day.tot=sum(individualCount,na.rm=T))%>%
  group_by(species,year)%>%
  mutate(prop=cumsum(day.tot/sum(day.tot,na.rm = T)))%>%
  filter(year>2004)

rh%>%
  ggplot(aes(j.day,prop))+geom_point()+facet_wrap(year~.)

rh.pred <- rh%>%
  group_by(year)%>%
  summarize(
    pred=predict(nls(prop~SSlogis(j.day,Asym, xmid, scal)),newdata=data.frame(j.day=min(j.day):max(j.day))), 
    j.day=min(j.day):max(j.day),
  )%>%
  left_join(rh%>%select(j.day,date)) 

rh%>%
  ggplot(aes(j.day,prop))+geom_point(aes=0.3)+geom_line(data=rh.pred,aes(x=j.day,y=pred),col="blue",size=2)+facet_wrap(year~.)

rh.arrive.date <-rh.pred%>%
  group_by(year)%>%
  filter(j.day==j.day[which.min(abs(pred-0.25))])

rh.arrive.date%>%
  ggplot(aes(year,j.day))+geom_point()

```
For Ruby-throated Hummingbird, the Julian day when 25% of the population arrived were 129-132.

```{r, "eBird data for belted kingfisher (bk)", warning=FALSE, results=FALSE}

bk<- dat.l%>%
  bind_rows(.id = "species") %>% 
  filter(species=="Megaceryle alcyon")%>%
  group_by(year)%>%
  mutate(date=as.Date(paste0(year,"-",month,"-",day)),
         j.day=julian(date,origin=as.Date(paste0(unique(year),"-01-01")))
  )%>%
  group_by(species,year,j.day,date)%>%
  summarise(day.tot=sum(individualCount,na.rm=T))%>%
  group_by(species,year)%>%
  mutate(prop=cumsum(day.tot/sum(day.tot,na.rm = T)))%>%
  filter(year>2004)

bk%>%
  ggplot(aes(j.day,prop))+geom_point()+facet_wrap(year~.)

bk.pred <- bk%>%
  group_by(year)%>%
  summarize(
    pred=predict(nls(prop~SSlogis(j.day,Asym, xmid, scal)),newdata=data.frame(j.day=min(j.day):max(j.day))), 
    j.day=min(j.day):max(j.day),
  )%>%
  left_join(bk%>%select(j.day,date)) 

bk%>%
  ggplot(aes(j.day,prop))+geom_point(aes=0.3)+geom_line(data=bk.pred,aes(x=j.day,y=pred),col="blue",size=2)+facet_wrap(year~.)

bk.arrive.date <-bk.pred%>%
  group_by(year)%>%
  filter(j.day==j.day[which.min(abs(pred-0.25))])

bk.arrive.date%>%
  ggplot(aes(year,j.day))+geom_point()

```
For Belted Kingfisher, the Julian day when 25% of the population arrived were 105-108.

```{r, "eBird data for yellow-bellied sapsucker (ys)", warning=FALSE, results=FALSE}
ys<- dat.l%>%
  bind_rows(.id = "species") %>% 
  filter(species=="Sphyrapicus varius")%>%
  group_by(year)%>%
  mutate(date=as.Date(paste0(year,"-",month,"-",day)),
         j.day=julian(date,origin=as.Date(paste0(unique(year),"-01-01")))
  )%>%
  group_by(species,year,j.day,date)%>%
  summarise(day.tot=sum(individualCount,na.rm=T))%>%
  group_by(species,year)%>%
  mutate(prop=cumsum(day.tot/sum(day.tot,na.rm = T)))%>%
  filter(year>2004)

ys%>%
  ggplot(aes(j.day,prop))+geom_point()+facet_wrap(year~.)

ys.pred <- ys%>%
  group_by(year)%>%
  summarize(
    pred=predict(nls(prop~SSlogis(j.day,Asym, xmid, scal)),newdata=data.frame(j.day=min(j.day):max(j.day))), 
    j.day=min(j.day):max(j.day),
  )%>%
  left_join(ys%>%select(j.day,date)) 

ys%>%
  ggplot(aes(j.day,prop))+geom_point(aes=0.3)+geom_line(data=ys.pred,aes(x=j.day,y=pred),col="blue",size=2)+facet_wrap(year~.)

ys.arrive.date <-ys.pred%>%
  group_by(year)%>%
  filter(j.day==j.day[which.min(abs(pred-0.25))])

ys.arrive.date%>%
  ggplot(aes(year,j.day))+geom_point()

```
For Yellow-bellied Sapsucker, the Julian day when 25% of the population arrived were 105-113.

```{r, "eBird data for common yellowthroat (cn)", warning=FALSE, results=FALSE}
#Common Yellowthroat,Geothlypis trichas(referred by cn)
cn<- dat.l%>%
  bind_rows(.id = "species") %>% 
  filter(species=="Geothlypis trichas")%>%
  group_by(year)%>%
  mutate(date=as.Date(paste0(year,"-",month,"-",day)),
         j.day=julian(date,origin=as.Date(paste0(unique(year),"-01-01")))
  )%>%
  group_by(species,year,j.day,date)%>%
  summarise(day.tot=sum(individualCount,na.rm=T))%>%
  group_by(species,year)%>%
  mutate(prop=cumsum(day.tot/sum(day.tot,na.rm = T)))%>%
  filter(year>2004)

cn%>%
  ggplot(aes(j.day,prop))+geom_point()+facet_wrap(year~.)


cn.pred <- cn%>%
  group_by(year)%>%
  summarize(
    pred=predict(nls(prop~SSlogis(j.day,Asym, xmid, scal)),newdata=data.frame(j.day=min(j.day):max(j.day))),#predict the logistic curve for each species
    j.day=min(j.day):max(j.day),
  )%>%
  left_join(cn%>%select(j.day,date))

cn%>%
  ggplot(aes(j.day,prop))+geom_point(aes=0.3)+geom_line(data=cn.pred,aes(x=j.day,y=pred),col="blue",size=2)+facet_wrap(year~.)

cn.arrive.date <-cn.pred%>%
  group_by(year)%>%
  filter(j.day==j.day[which.min(abs(pred-0.25))])

cn.arrive.date%>%
  ggplot(aes(year,j.day))+geom_point()
```
For Common Yellowthroat, the Julian day when 25% of the population arrived were 131-133.

Overall, depending on the species, it seems like the average arrival timings were either around 105-110 or 128-133.
```{r, "weather data", echo=TRUE, include=FALSE, results=FALSE, message=FALSE}
weather.d <- weather.d%>%
  dplyr::mutate(year=as.integer(str_sub(date,1,4)), 
                date=as.Date(date))%>%
  group_by(year)%>% 
  dplyr::mutate(j.day=julian(date,origin=as.Date(paste0(unique(year),"-01-01"))), 
                date2=date,
                wdir.rad=(180-abs(wdf2-180))*pi/180, 
                wvec=cos(wdir.rad)*-1*awnd 
  )%>% 
  select(id,year,date2,j.day,tmin,tmax,wvec)%>% 
  left_join(sta.d%>%select(id,name,migr.day))%>% 
  mutate(j.day=j.day+migr.day)
```

```{r, "mean arrival time for cs", include=FALSE, message=FALSE, result=FALSE}
cs.arr.weath <- cs.arrive.date%>%
  left_join(weather.d)%>%
  left_join(cs%>%select(year,date,j.day))

head(cs.arr.weath)

weather.wk <-weather.d %>% 
  group_by(year,name) %>% 
  mutate(wk.tmin = frollmean(tmin, n=14,align="right"),
         wk.tmax = frollmean(tmax, n=14,align="right"),
         wk.wvec = frollmean(wvec, n=14,align="right")
  )%>%
  select(j.day,date2,name,wk.tmin,wk.tmax,wk.wvec)

cs.arr.weath2 <- cs.arrive.date%>%
  left_join(weather.wk)

head(cs.arr.weath2)
```

```{r, "mean arrival time for rh", include=FALSE, message=FALSE, result=FALSE}
rh.arr.weath <- rh.arrive.date%>%
  left_join(weather.d)%>%
  left_join(rh%>%select(year,date,j.day))

head(rh.arr.weath)

weather.wk <-weather.d %>% 
  group_by(year,name) %>% 
  mutate(wk.tmin = frollmean(tmin, n=14,align="right"),
         wk.tmax = frollmean(tmax, n=14,align="right"),
         wk.wvec = frollmean(wvec, n=14,align="right")
  )%>%
  select(j.day,date2,name,wk.tmin,wk.tmax,wk.wvec)

rh.arr.weath2 <- rh.arrive.date%>%
  left_join(weather.wk)

head(rh.arr.weath2)
```

```{r, "mean arrival time for bk", include=FALSE, message=FALSE, result=FALSE}
bk.arr.weath <- bk.arrive.date%>%
  left_join(weather.d)%>%
  left_join(bk%>%select(year,date,j.day))

head(bk.arr.weath)

weather.wk <-weather.d %>% 
  group_by(year,name) %>% 
  mutate(wk.tmin = frollmean(tmin, n=14,align="right"),
         wk.tmax = frollmean(tmax, n=14,align="right"),
         wk.wvec = frollmean(wvec, n=14,align="right")
  )%>%
  select(j.day,date2,name,wk.tmin,wk.tmax,wk.wvec)

bk.arr.weath2 <- bk.arrive.date%>%
  left_join(weather.wk)

head(bk.arr.weath2)
```

```{r, "mean arrival time for ys", include=FALSE, message=FALSE, result=FALSE}
ys.arr.weath <- ys.arrive.date%>%
  left_join(weather.d)%>%
  left_join(ys%>%select(year,date,j.day))

head(ys.arr.weath)

weather.wk <-weather.d %>% 
  group_by(year,name) %>% 
  mutate(wk.tmin = frollmean(tmin, n=14,align="right"),
         wk.tmax = frollmean(tmax, n=14,align="right"),
         wk.wvec = frollmean(wvec, n=14,align="right")
  )%>%
  select(j.day,date2,name,wk.tmin,wk.tmax,wk.wvec)

ys.arr.weath2 <- ys.arrive.date%>%
  left_join(weather.wk)

head(ys.arr.weath2)
```

```{r, "mean arrival time for cn", include=FALSE, message=FALSE, result=FALSE}
cn.arr.weath <- cn.arrive.date%>%
  left_join(weather.d)%>%
  left_join(cs%>%select(year,date,j.day))

head(cn.arr.weath)

weather.wk <-weather.d %>% 
  group_by(year,name) %>% 
  mutate(wk.tmin = frollmean(tmin, n=14,align="right"),
         wk.tmax = frollmean(tmax, n=14,align="right"),
         wk.wvec = frollmean(wvec, n=14,align="right")
  )%>%
  select(j.day,date2,name,wk.tmin,wk.tmax,wk.wvec)

cn.arr.weath2 <- cn.arrive.date%>%
  left_join(weather.wk)

head(cn.arr.weath2)
```

```{r, "linear mixed-effect modeling for cs species", results=FALSE, message=FALSE, warning=FALSE}
cs.lmer <- lmer(j.day~tmin*tmax*wvec+(1|name),cs.arr.weath,na.action = "na.fail")
Anova(cs.lmer)
cs.lmer2 <- lmer(j.day~wk.tmin*wk.tmax*wk.wvec+(1|name),cs.arr.weath2,na.action = "na.fail")
Anova(cs.lmer2)


cs.arr.aic <- dredge(cs.lmer2,fixed = c("wk.tmin","wk.tmax","wk.wvec"),)

cs.kb <- kable(cs.arr.aic[1:4,],caption = "Fit values for nested models of the most complicated lme model")
kable_styling(cs.kb)

best.lmer <-  lmer(j.day~wk.tmin+wk.tmax+wk.wvec+(1|name),cs.arr.weath2,na.action = "na.fail")

Anova(best.lmer)
```
Based on the collection of weather and bird data that underwent data analysis, a linear mixed-effect modeling for each of the species was conducted. For the Chimney swift species, it was found that tmin and wind velocity are significant. Through the best model ANOVA analysis, wind velocity is a significant variable.

```{r, "linear mixed-effect modeling for rh species", results=FALSE, message=FALSE, warning=FALSE}
rh.lmer <- lmer(j.day~tmin*tmax*wvec+(1|name),rh.arr.weath,na.action = "na.fail")
Anova(rh.lmer)
rh.lmer2 <- lmer(j.day~wk.tmin*wk.tmax*wk.wvec+(1|name),rh.arr.weath2,na.action = "na.fail")
Anova(rh.lmer2)


rh.arr.aic <- dredge(rh.lmer2,fixed = c("wk.tmin","wk.tmax","wk.wvec"),)

rh.kb <- kable(rh.arr.aic[1:4,],caption = "Fit values for nested models of the most complicated lme model")
kable_styling(rh.kb)

best.lmer <-  lmer(j.day~wk.tmin+wk.tmax+wk.wvec+(1|name),rh.arr.weath2,na.action = "na.fail")

Anova(best.lmer)

```
For the Ruby-throated Hummingbird species, it was found that tmin and wind velocity are significant. Through the best model ANOVA analysis, wind velocity was the most significant variable.

```{r, "linear mixed-effect modeling for bk species", results=FALSE, message=FALSE, warning=FALSE}

bk.lmer <- lmer(j.day~tmin*tmax*wvec+(1|name),bk.arr.weath,na.action = "na.fail")
Anova(bk.lmer)
bk.lmer2 <- lmer(j.day~wk.tmin*wk.tmax*wk.wvec+(1|name),bk.arr.weath2,na.action = "na.fail")
Anova(bk.lmer2)


bk.arr.aic <- dredge(bk.lmer2,fixed = c("wk.tmin","wk.tmax","wk.wvec"),)

bk.kb <- kable(bk.arr.aic[1:4,],caption = "Fit values for nested models of the most complicated lme model")
kable_styling(bk.kb)

best.lmer <-  lmer(j.day~wk.tmin+wk.tmax+wk.wvec+(1|name),bk.arr.weath2,na.action = "na.fail")

Anova(best.lmer)
```
For the Belted-Kingfisher species, it was found that wind velocity is significant. Through the best model ANOVA analysis, wind velocity and tmin are significant variables.

```{r, "linear mixed-effect modeling for ys species", results=FALSE, message=FALSE, warning=FALSE}
ys.lmer <- lmer(j.day~tmin*tmax*wvec+(1|name),ys.arr.weath,na.action = "na.fail")
Anova(ys.lmer)
ys.lmer2 <- lmer(j.day~wk.tmin*wk.tmax*wk.wvec+(1|name),ys.arr.weath2,na.action = "na.fail")
Anova(ys.lmer2)


ys.arr.aic <- dredge(ys.lmer2,fixed = c("wk.tmin","wk.tmax","wk.wvec"),)

ys.kb <- kable(ys.arr.aic[1:4,],caption = "Fit values for nested models of the most complicated lme model")
kable_styling(ys.kb)


best.lmer <-  lmer(j.day~wk.tmin+wk.tmax+wk.wvec+(1|name),ys.arr.weath2,na.action = "na.fail")

Anova(best.lmer)
```
For the Yellow-bellied Sapsucker species, it was found that there is no significance in any variable. Through the best model ANOVA analysis, there's no specific significance for any variable.

```{r, "linear mixed-effect modeling for ys species", results=FALSE, message=FALSE, warning=FALSE}
cn.lmer <- lmer(j.day~tmin*tmax*wvec+(1|name),cn.arr.weath,na.action = "na.fail")
Anova(cn.lmer)
cn.lmer2 <- lmer(j.day~wk.tmin*wk.tmax*wk.wvec+(1|name),cn.arr.weath2,na.action = "na.fail")
Anova(cn.lmer2) 

cn.arr.aic <- dredge(cn.lmer2,fixed = c("wk.tmin","wk.tmax","wk.wvec"),)

cn.kb <- kable(cn.arr.aic[1:4,],caption = "Fit values for nested models of the most complicated lme model")
kable_styling(cn.kb)

best.lmer <-  lmer(j.day~wk.tmin+wk.tmax+wk.wvec+(1|name),cn.arr.weath2,na.action = "na.fail")

Anova(best.lmer)
```
For the Common Yellowthroat species, it was found that wind velocity is significant along with both variables tmin and tmax together. Through the best model ANOVA analysis, wind velocity and tmin are significant variables.

### Discussion


### Contributions
Steven worked on R-coding, the Methods write-up, and data analysis. Nick worked on the Introduction, Results, and Discussion sections. Tony helped with problem solving during the coding process. Amanda helped with the coding portion of the project along with the Results and References sections. Amanda and Steven worked on data analysis. All team members carried out editing at the end.

### References